---
title: "Practice Midterm Exam 1"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 3: Data Analysis

For this part, you will need to use R and the following dataset: 

The dataset was collected from Airbnb with data on listings in the city of Asheville, NC, and can be downloaded here: https://gatech.instructure.com/courses/139200/files/folder/Practice%20Midterm%201?preview=15402865. 

Here is the data provided for each listing:

room id: A unique number identifying an Airbnb listing.
host id: A unique number identifying an Airbnb host.
room type: One of 'Entire home/apt','Private room', or 'Shared room'
reviews: The number of reviews that a listing has received.
overall satisfaction: The average rating (out of five) that the listing has received from those vistors who left a review.
accommodates: The number of gusts a listing can accommodate.
bedrooms: The number of bedrooms a listing offers.
price: The price (in USD) for a night stay. In early surveys, there may be some values that were recorded by month.
Fit a multiple linear regression model named 'model1' using price as the response variable and the following predicting variables: room type, reviews, overall satisfaction, accommodates, and bedrooms.

To read the data in R, save the file in your working directory (make sure you have changed the directory if different from the R working directory) and read the data using the R function read.csv().

```{r}
# Read in the data
data = read.csv("tomslee_airbnb_asheville_1498_2017-07-20.csv", head = TRUE, sep = ",")
# Show the first few rows of data
head(data, 3)
```

## Question 25

Which coefficients (including intercept) are statistically significant at the 99% confidence level?

```{r}
# Your code here...

```

## Question 26

What is the estimated coefficient for room type = "Private Room"?

```{r}
# Your code here...

```

## Question 27

What is the interpretation for the estimated coefficient for room type = "Private Room"?

```{r}
# Your code here...

```

## Question 28

Assess whether the model assumptions hold.

```{r}
# Your code here...

```

## Question 29

Using your model, make a prediction for a listing on Airbnb in Asheville with the following factors: bedrooms = 1, accommodates = 2, reviews = 92, overall_satisfaction = 3.5, and room_type = 'Private room'. What is your predicted price for such a listing and the corresponding 95% prediction interval?

```{r}
# Your code here...

```
