---
title: "Practice Midterm 2"
author: "ISYE 6414 OAN"
date: "2020 Fall"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data analyzed in this part of the exam came from medical patients who have suffered from a heart attack within the past year. The data can be downloaded at https://gatech.instructure.com/courses/139200/files/13984885/download?wrap=1.

The response variable is *still_alive* which is a binary variable indicating if the patient is still living or not (0=dead, 1=alive).

The predicting variables are as follows:

* age -- age in years when heart attack occurred
* pericardial_effusion-- Pericardial effusion is fluid around the heart. This is a binary variable: 0=no fluid, 1=fluid.
* fractional_shortening -- a measure of contractility around the heart. Lower numbers are increasingly abnormal
* lvdd -- left ventricular end-diastolic dimension. This is a measure of the size of the heart at end-diastole. Large hearts tend to be sick hearts


Read the data and answer the questions below:
```{r}
# Loading of the data

echo= read.csv("echo.csv", header=TRUE, sep=",")

# Viewing first few rows of data

head(echo)
```


## Q21
Plot the boxplots to describe the relationship between the response (still_alive) and the following predicting variables: *age*,  *fractional_shortening*, and *lvdd*. Intepret the plots.

```{r}
## Your code here...

```

## Q22
Create a logistic regression model using the following predicting variables: *age*, *fractional_shortening*, and *lvdd*.  

a) What are the model parameters and what are their estimates? 

b) Which of the parameters are statistically significant at the 95% confidence level? 

```{r}
## Your code here...

```

## Q23
Provide the equation for the estimated logit transformation of the probability of still being alive given the predicting variables. 



## Q24
Interpret the estimated value of the parameter corresponding to *age* in the context of this problem. 



## Q25
Perform a test for the overall regression, using $\alpha = 0.05$. Does the overall regression have explanatory power?
```{r}
## Your code here...

```

## Q26
How does the log-odds change with a 1 unit increase in *fractional_shortening*, holding all other variables constant? Explain.



## Q27
What is the approximated distribution of the null deviance? What is the value of the null deviance and its degree of freedom?

